A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-06-01 manuscript.Rmd
reinclude chunk for container size
explain what CLI is and how we make it easy closes #112

First, to resolve ambiguity across multiple versions of code and data that may be come into existence during the development phase of a research project, we recommend the use of a version control system.
A version control system tracks changes to all project-related files (e.g., materials, data, and code) over time.
At a later stage, single files or the entire project can be compared to or reverted to an earlier version.
Additional features such as branches allow parallel versions to exist that can be merged later.
Specific snapshots of a repository can be marked, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
Using version control protects against loss of files and also supports remote collaboration with semi-automated tools for tracking and merging changes.

@@

Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
This paradigm has gained popularity with the Sweave approach [@leisch2002sweave] that allowed embedding the data analysis into a TeX document, which then could be dynamically rendered into a final manuscript.
As of recently, R Markdown has added to the success of this idea.
Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
This way, Markdown does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles] and presentations[@revealjs]---and even your CV [@vitae].
R Markdown extends regular Markdown by allowing users to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
Upon rendering of the document, the code blocks are executed and their output is dynamically inserted into the document.
This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).

One would hope that sharing a literate statistical analysis would allow others (including one own's future self) to simply download and run (ie., reproduce) a given analysis.
However, as @ argue, in most instances this approach fails.
In practice, dependencies on external software (such as system libraries or components of the programming language) are frequently unmentioned and the components or packages are not included.
Troubleshooting  issues specific to a certain programming language or dependent tool typically requires considerable expertise and hinder reproducibility.
One solution to this problem is to provide not only the analysis script but all dependent software packages and system libraries in a so-called _container_.
"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silver2017].
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called an "image".
Packaging all needed software in such an image requires considerable amounts of storage space.
There are two major strategies helping to keep the storage requirements reasonable.
First, there is a community that maintains images for particular purposes.
For example, there are ready-made images for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`

@@

@@

A preregistration as code offers four advantages over classical preregistration.
First, a PAC removes any ambiguity regarding the translation of an analysis plan into code.
Second, despite being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
Researchers can, for example, formulate conditions under which they prefer one analysis over the other, i.e. if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
Third, deviations from the preregistration are more explicit because they are reflected in changes to the code a property facilitated by version control.
Fourth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.

>>>>

First, to resolve ambiguity across multiple versions of code and data that may be come into existence during the development phase of a research project, we recommend the use of a version control system.
A version control system tracks changes to all project-related files (e.g., materials, data, and code) over time.
At a later stage, single files or the entire project can be compared to or reverted to an earlier version.
Additional features, such as branches, facilitate collaboration.
Specific snapshots of a repository can be marked, such that one can clearly label a given project status as the one that created a given preregistration, a given preprint, or the final authors' version accepted at a journal.
Using version control protects against loss of files and also supports remote collaboration with semi-automated tools for tracking and merging changes.

@@

Dynamic document generation fuses both steps and can be traced back to @knuthCWEBSystemStructured who put it this way: "The main idea is to regard a program as a communication to human beings rather than as a set of instructions to a computer."
This paradigm has gained popularity with the Sweave approach [@leisch2002sweave] that allowed embedding the data analysis into a TeX document, which then could be dynamically rendered into a final manuscript.
As of recently, R Markdown has added to the success of this idea.
R Markdown uses Markdown for writing the text, and R (or other programming languages) for the analysis. Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
This way, Markdown does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX), works well with version control systems, and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles] and presentations[@revealjs]---and even your CV [@vitae].
R Markdown extends regular Markdown by allowing users to include R code chunks (in fact, arbitrary computer code; see @riedererChapter15Other) into a Markdown document.
Upon rendering of the document, the code blocks are executed and their output is dynamically inserted into the document.
This allows the creation of (conditionally) formatted text, statistical results, or figures that are guaranteed to be up-to-date because they are created every time anew as the document is rendered to its output format (e.g., presentation slides or a journal article).

One would hope that sharing a literate statistical analysis would allow others (including one own's future self) to simply download and run (ie., reproduce) a given analysis.
However, as @ argue, in most instances this approach fails.
In practice, dependencies on external software (such as system libraries or components of the programming language) are frequently unmentioned and the components or packages are not included.
Troubleshooting  issues specific to a certain programming language or dependent tool typically requires considerable expertise and hinder reproducibility.

```{r, include=FALSE}
container_size <- as.numeric(fs::file_size("reprotutorial.sif"))/1024^3
```

One solution to this problem is to provide not only the analysis script but all dependent software packages and system libraries in a so-called _container_.
"By packaging the key elements of the computational  environment  needed  to  run  the  desired  software, [...] they  make the software much easier to use, and the results easier to reproduce [@silver2017].
Instead of leaving it to the user which software should be installed how on their computer, Docker provides all software used in an analysis, including the operating system itself.
The operating system level is important because some functionality may be passed to software layers beneath the programming language, such as calls to random number generators, linear algebra libraries and such.
Docker does this without interfering with the already installed software by using a virtual software environment that is independent of the host software environment.
Such a snapshot of the software stack is called an "image".
Packaging all needed software in such an image requires considerable amounts of storage space.
There are two major strategies helping to keep the storage requirements reasonable.
First, there is a community that maintains images for particular purposes.
For example, there are ready-made images for R, based on Ubuntu (a Linux operating system) containers [@boettigerIntroductionRockerDocker2017].
Users can then install whatever they need in addition to what is provided by these pre-compiled images.
`r ifelse(is.na(container_size), "", glue::glue("The image that was used for this article uses {round(container_size, 2)}Gb of disc space."))`
Our image includes Ubuntu, R, R Studio, LaTeX as well as a variety of R packages like the tidyverse [@tidyverse], amounting to `r installed.packages() %>% as.data.frame() %>% filter(is.na(Priority)) %>% nrow()` packages.
Second, instead of saving a binary image of the software, one can also save a container recipe, a so-called  `Dockerfile`.
This recipe is a textual description of all commands that need to be executed to recreate the image.
Such files are tiny (the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/Dockerfile", "Dockerfile")` for this project is `r round(as.numeric(fs::file_size("Dockerfile"))/1024^1, 2)`Kb big) and are an unambiguous description of how to recreate a given container image.
However, they rely on the assumption that all software repositories will remain accessible and will continue to make available historic software versions.
For archiving, we therefore recommend that a full image is always kept (see Section on [Archiving Docker]).
With the `repro`-package adding R packages dependencies is as simple as adding them to the header under the tag `packages:`

@@

A major hurdle for adoption of the before mentioned tools for researchers is that they are typically used through a command line interface (CLI).
The CLI is vastly more user friendly than its predecessor, the punch card, but many users (including the authors) prefer a graphical user interface (GUI) over text based interactions for most tasks.
A GUI has the obvious disadvantages for reproducibility, that it requires a user who clicks on the correct buttons in the correct order.
Depending on the number of steps this is not only much effort but also error prone.
This tutorial seeks a compromise.
We use a GUI (and similar helpers) where we can, to create a reproducible workflow but the reproduction of such workflow does not rely on manual steps.
At several points we rely on or have self developed a CLI that wraps a more complicated (and powerful) CLI.
These wrapper use standard CLI tools but provide feedback for the user about what the computer does and what the user should do in laymans terms.
We hope this make reproducibility tools more accessible by enabling untrained users to detect their systems state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
These alternative tools are merely an assistance system and as users get more comfortable they can use the underlying tools directly to solve more complex problems.

One example for such assistance are the check functions from `repro`.

Two properties of CLI tools make them useful but at the same time difficult to learn.
Many CLI tools are decades old, and are therefore stable, widely used, and widely supported.
Because of their age they seem archaic and are difficult to use when compared to more modern computer tools.
These tools are also more powerful than many GUI counterparts, because of their flexibility---many options to change their behavior and composabibility---the output of one, can be the input of another.

@@

A preregistration as code offers four advantages over classical preregistration.
First, a PAC removes any ambiguity regarding the translation of an analysis plan into code.
Second, despite being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
Researchers can, for example, formulate conditions under which they prefer one analysis over the other, i.e. if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
Third, deviations from the preregistration are more explicit because they are reflected in changes to the codes a property facilitated by version control.
Fourth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.
